The Power of Two Random Choices\\ \large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Tr\"aff and Francesco Versaci, lecture ``Parallel Algorithms'', 2012 winter term at TU Wien.}

### 2014-02-21 article/article.tex
continued fixing stuff

\subsection{Multiple-choice BiB}
Consider a variation of the above single-choice BiB strategy: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin. If there are multiple bins with the same minimal load, chose one among them uniformly and independently at random. This simple multiple-choice scheme, which we will call the Greedy algorithm, achieves a gap
\begin{align*}
\delta = \frac{\log \log n}{\log d} + O(1)
\end{align*}
 w.h.p.~for all $m$~\cite{ABKU99}~\cite{BCSV06}.

This apparently minor change introduces a small amount of choice which, compared to the single-choice scheme, leads to a significant reduction in the maximum load. 
\begin{comment}
Curiously, the exponential decrease in the gap is already achieved having just two choices (that is, $d=2$). Each additional choice decreases the bound for the maximum gap by just a constant factor~\cite{MRS01}. 
\end{comment}
Extensive research following the influential paper by Azar et al.~\cite{ABKU99}, who were the first to explicitly prove this phenomenon\footnote{For a brief history of the proof, which is quite interesting by itself, refer to Appendix \ref{sec:historyOfGreedy}.}, suggests that this two-choice paradigm is not just an artifact of the simple BiB model, but a general principle applicable to a variety of problems.

Consider a slightly more involved strategy, which we will call the Always-go-left scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size\footnote{For the sake of a clear presentation, we will assume that $d$ divides $n$ evenly in the remainder of this paper.} $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$. We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins. In case a tie occurs, put the ball into the left-most bin. That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with the same minimal load, we put the ball into the bin with smallest index $i$. Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection reduces the bound for the gap to\footnote{The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers. Refer to Appendix \ref{sec:fibonacci} for the definition.}
\begin{align*}
\delta = \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
\end{align*}
w.h.p. Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants \cite{VOC03}.

This paper is organized as follows: We give a basic classification of BiB games in Section \ref{sec:classification}, to provide some overview of the different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}. In the remaining sections of this paper we will consider the model given in Section \ref{sec:model}, which can be classified as a sequential, finite and dynamic process with random deletions. Section \ref{sec:multiple-choice} starts with a brief definition of three different classes of random placement strategies and then continues with the presentation of the two placement algorithms outlines in the introduction: The  Greedy algorithm (Section \ref{sec:greedy}) and the Always-go-left algorithm (Section \ref{sec:AlwaysGoLeft}). Section \ref{sec:analysis} is the core of this paper and will give a detailed analysis of both strategies. We will use a proof technique called witness trees, which allows for a clean analysis of both problems. Note that the technique is applicable to a wide variety of BiB scenarios, making it the most useful of several proof techniques applied to BiB problems. Other proof techniques such as layered induction and a differential equations approach have been used. They are however not part of this paper. See, e.g., \cite{MRS01} for an overview of different proof techniques for the BiB problem.

@@

\section{Classification}
\label{sec:classification}
Several different variations of this problem exist, which may be classified by the following properties:
\begin{compactitem}
\item In a \emph{sequential} process, the balls arrive one after another. A  ball is dealt with only after the previous ball has been placed. Contrary, in a \emph{parallel} process, balls arrive in sets and the balls in one set are dealt with simultaneously \cite{ABS98}. 
\item In a \emph{static} problem, the balls are placed into the bins and no balls are ever removed. In a \emph{dynamic} problem, balls can also be removed from the system. In a simple random deletions model, for each newly inserted ball a ball chosen uniformly and independently at random from among the balls currently in the bins is removed from the system \cite{ABKU99} \cite{MRS01}. More elaborate schemes exists: An oblivious adversary specifies the sequence of insertions and deletions of balls in advance, without any knowledge of the placement of the balls \cite{CFM+98}. Another approach was studied by Adler et al.~\cite{ABS98} in the context of the parallel BiB problem: In each bin, the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted. 
\item A \emph{finite} process is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite} process, a placement and a deletion process is defined over an infinite time span \cite{ABS98}.
\item In the lightly loaded case, $m$ is in the order of $n$, whereas in the heavily loaded case, $m$ is arbitrarily large. For clarity, we will assume that $m = n$ in the lightly loaded case and that $m > n$ in the heavily loaded case. 
\end{compactitem}

\section{Applications}
\label{sec:applications}
The two-choice paradigm applied to the BiB model has several useful applications, some of which we will outline here. The applications in Hashing (Section \ref{sec:hashing}), Online Load Balancing (Section \ref{sec:loadbalancing}) and Emulation of shared memory machines on distributed memory machines \ref{sec:DMM}) share a close relation to load balancing. Section \ref{sec:circuitrouting} applies the BiB model and the two-choice paradigm in a different context, namely that of routing circuits in multi-stage interconnection networks with minimal congestion.

@@

It is easy to see how the BiB idea can be applied: Assuming a perfect hash function\footnote{i.e., each of the $n$ elements is mapped to one of the $n$ table entries uniformly and independently at random}, the elements are balls to be mapped to table entries or bins, implying a worst case look-up time of $\Theta\left(\frac{\log n}{\log \log n}\right)$ w.h.p. Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries implies a bound of $\Theta\left(\log \log n\right)$ w.h.p. This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions \cite{ABKU99} \cite{MRS01}.

\subsection{Online load balancing}
\label{sec:loadbalancing}
The following simple scenario is a typical application of online load balancing: Consider $n$ servers\footnote{Such as file-, database- or network servers} and $m$ requests issued by clients that need to be handled by these servers. For simplicity, assume that each task can be handled by any server and that all tasks are of the same size, i.e., require the same amount of execution time.

To maximize the throughput of the servers, a uniform distribution of tasks to servers is desirable. This can easily be achieved by a central dispatcher. However, such a central dispatcher poses a bottleneck within a distributed system and querying $m$ servers for their current load involves significant overhead, e.g., sending a message to each server and waiting for the reply and more efficient solutions are necessary. As with hashing, the BiB model can be applied easily: we need to assign tasks (the balls) to servers (the bins). Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends issues its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is bounded by $\Theta\left(\log \log n\right)$ w.h.p. \cite{KLM92} \cite{RS98} \cite{MRS01}.

\subsection{Emulation of shared memory machines on distributed memory machines}
\label{sec:DMM}
The Parallel Random Access Machine model\footnote{as defined in the lecture or in~\cite{P03}}, or PRAM for short, is a high level abstraction of a shared memory machine, in which memory access by any processor to any memory location is assumed to require constant time. It is especially useful for the study of parallel algorithms. However, a PRAM is much harder to realize in hardware than a Distributed Memory Machine (DMM). Thus, ways to emulate a PRAM on a DMM have been studied extensively. Such an emulation distributes the processors and memory cells of the PRAM to the processors and memory modules (typically one per processor) of the DMM with the goal to reduce the slowdown of the emulation, which is the time the DMM requires to simulate one step of the PRAM. Each memory module can handle only one memory access at a time step. More accesses to the same module will result in memory contention. The BiB idea is useful for managing communication between the processors and the memory modules such that memory contention is minimal. For details, see Karp et al.~\cite{KLM92} and the papers referenced there.

\subsection{Low congestion circuit routing}
\label{sec:circuitrouting}
In multi-stage interconnection networks, virtual circuit-switching is used to route communication. For simplicity, we consider the permutation routing problem, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and assume a butterfly network as the underlying interconnection network (see \cite{CLR09} for a detailed definition). Then, congestion is the maximum number of paths that are routed through a network link. The goal of a circuit routing algorithm is to to allocate paths for all $n$ communication requests such that congestion is minimal.

@@

Applying the two-choice paradigm in this context is somewhat more involved and beyond the scope of this paper. The interested reader may find more details in \cite{CMM+98}.

@@

Consider a variation of the above one-choice BiB strategy, which we will call the Greedy scheme: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin. If there are multiple bins with the same minimal load, chose one among them arbitrarily. Note that the $d$ selected bins are not necessarily distinct.

\begin{theorem}
\label{theorem:greedy}
Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is at most 
\begin{align*}
\frac{m}{n} + \frac{\log \log n}{\log d}+ O(h)
\end{align*}
w.h.p.~at any time $t$ \cite{ABKU99} \cite{BCSV06}.
\end{theorem}

This implies a gap of $\Theta\left(\log \log n \right)$ for arbitrarily large $m$, which is remarkable in several ways:
\begin{compactitem}
\item The apparently small change made for the Greedy algorithm compared to the single-choice BiB scheme results in an exponential decrease of the gap, even for $d=2$. 
\item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{MRS01}. 
\item For the heavily loaded case, the resulting gap of the multiple-choice BiB strategy does not depend on the number of balls $m$. In contrast, the bound for the one-choice BiB scheme diverges with $m$.
\begin{comment}
\todo{the following is only true for class 1 algorithms}
\item The given bounds are tight, meaning that no other strategy that places each ball into one of $d$ randomly selected bins achieves a gap that is asymptotically lower.
\end{comment}
\end{compactitem}

@@

\begin{theorem}
\label{theorem:agln}
If $n$ balls are placed into $n$ bins using the Always-go-left scheme, the number of balls in the fullest bin will be 
\begin{align*}
\frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
\end{align*}
w.h.p.
\end{theorem}

\begin{theorem}
\label{theorem:algm}
Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, at any time $t$ the maximum load is at most 
\begin{align*}
\frac{\ln\ln n}{d  \ln \Phi_d} + O(h)
\end{align*}
w.h.p.
\end{theorem}

\section{Analysis}
\label{sec:analysis}
In this section, we will prove Theorems \ref{theorem:greedy}, \ref{theorem:agln} and \ref{theorem:algm} in detail, using witness trees to provide an upper bound for the event that a bin contains too many balls. Our analysis closely follows the proofs in \cite{VOC03}. \todo{is it enough to cite this once for the whole section?}

@@

We define edge and leaf events:
\begin{compactitem}
\item \emph{Edge event:} Consider some edge $e = (u,v)$ of the witness tree, s.t.~$v$ is the i-th child of $u$. Then $e$ represents the event that the i-th location of the ball $b_u = \mathrm{ball}(u)$ points to the same bin as one of the locations of the ball $b_v = \mathrm{ball}(v)$. 
%$\todo{illustrate}
\item \emph{Leaf event:} A leaf node $v$ of the witness tree represents the event that each of the $d$ locations of the ball $b_v = \mathrm{ball}(v)$ points to a bin that contains at least three additional balls (i.e., balls that are not represented by a node of the tree) at the time of the insertion of the ball $b_v$.
%\todo{illustrate}
\end{compactitem}
The definition for leaf events refers to bins containing some number of balls at some specified time. To avoid dependencies, the crucial trick is that edge events are defined in terms of the alternative locations of balls, not by their finally assigned bin.

@@

An edge\footnote{where $v$ is the i-th child of $u$} $(u, v)$ is activated with probability at most $d/n$, which is the probability that the i-th location of the ball $b_u = \mathrm{ball}(u)$ points to a particular location of the ball $b_v = \mathrm{ball}(v)$ times the number of locations $d$. This implies a probability of at most
\begin{align*}
p_e = \left(\frac{d}{n}\right)^{k-1}
\end{align*} 
or the event that all $k-1$ edges of a witness tree are activated. Note that  the assumption that all balls are distinct (Assumption \ref{assumption:independence}) implies that all edge events are independent.

@@

Let $x$ be the bin that contains $L+4$ balls. The root node gets assigned the topmost ball in bin $x$. Each of the $d$ locations of this ball points to a bin with at least $L+3$ balls in it, at the time of its insertion. These are the $d$ bins that were sampled by the Always-go-left scheme and thus bin $i$ belongs to the $i$'th group of bins, where each group consists of $n/d$ bins\footnote{For the sake of simplicity, we assume that $d$ divides $n$ evenly, i.e., that all groups have the exact same size. It can be shown that the analysis still holds if each group is of size $\lfloor\frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$.}. The child node of the root with label $(L-1, i)$ is assigned the topmost ball of the bin the $i$' location points to. The assignment of balls to nodes than proceeds recursively. Given a node $v$ with label $(h, i)$ whose ball $b_v$ was put into some bin belonging to group $i$ and containing at least $l+3$ other balls, we can conclude the following.
\begin{compactitem}
\item For $1\leq j < i$, the $j$'th location of the ball $b_v$ points to a bin containing at least $l+4$ balls at the insertion time of the ball $b_v$. If the j-th location contained less balls than the i-th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$. The topmost ball in location $j$ is assigned to the  child $u$ of $v$ with label $(l,j)$. 
\item For $i < j \leq d$, the j-th location contains at least $l+3$ balls. If location $j$ contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin it points to. The topmost ball in the bin that location $j$ points to is assigned to the child node $u$ of $v$ labeled with $(l-1, j)$. \todo{the paper claims that the locations $j$ point to bins with $l+2$ balls. This can not be true, since the topmost, i.e., the $l+2$ ball - call it $b_t$ - gets assigned to a node with label $(l'=l-1, j)$. Since there are $l+1=l'+2$ balls below it, at the time of the placement of ball $b_t$, the bin $j$ it gets put into has only $l'+2$ balls in it, violating the invariant. Same with the case $j<i$.} 
\end{compactitem}
\end{proof}

@@

\subsubsection{Construction}
\label{sec:analysis:constructionFullWT}
Assume a bin $x$ contains at least $L+3+\gamma$ balls at time $t$. We construct a witness tree as follows.
\begin{compactitem}
\item The root is not assigned any ball.
\item Each child of the root is assigned one of the topmost $\gamma$ balls in bin $x$, which are necessarily distinct.
\item For the assignment of nodes to the grandchildren of the root, consider a ball $b_c$ that is assigned to a child node $v_c$. Let the child node of $v_c$ be denoted by $v_g$. Since the ball $b$ was placed into bin $x$, at least one of its locations points to this bin $x$, which we call location $i$. Let $j = i+1 mod d$ be the ball's j-th location. Then, at the time of insertion of ball $b_c$, its j-th location points to a bin $y$ with at least $L+3$ other balls in it. We assign the topmost ball in bin $y$ to node $v_g$.
\end{compactitem}
Notice that the $\gamma$ additional balls in bin $x$ are assigned to the child nodes of the root only.

>>>>

\subsection{Multiple-choice BiB}
Consider a variation of the  single-choice BiB strategy: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin. If there are multiple bins with the same minimal load, chose one among them uniformly and independently at random. This simple multiple-choice scheme, which we will call the Greedy algorithm, achieves a gap
\begin{align*}
\delta = \frac{\log \log n}{\log d} + O(1)
\end{align*}
 w.h.p.~for all $m$~\cite{ABKU99}~\cite{BCSV06}.

This apparently minor change introduces a small amount of choice which leads to a significant reduction in the maximum load. 
\begin{comment}
Curiously, the exponential decrease in the gap is already achieved having just two choices (that is, $d=2$). Each additional choice decreases the bound for the maximum gap by just a constant factor~\cite{MRS01}. 
\end{comment}
Extensive research following the influential paper by Azar et al.~\cite{ABKU99}, who were the first ones to explicitly prove this phenomenon\footnote{For a brief history of the proof, which is quite interesting by itself, refer to Appendix \ref{sec:historyOfGreedy}.}, suggests that this two-choice paradigm is not just an artifact of the simple BiB model, but a general principle applicable to a variety of problems.

Consider the slightly more Always-go-left scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size\footnote{To simplify the presentation, we will assume that $d$ divides $n$ evenly in the remainder of this paper.} $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$. We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins. In case a tie occurs, we put the ball into the left-most bin. That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with the same minimal load, we put the ball into the bin with smallest index $i$. Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection reduces the bound for the gap to\footnote{The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers. Refer to Appendix \ref{sec:fibonacci} for the definition.}
\begin{align*}
\delta = \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
\end{align*}
w.h.p. Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants \cite{VOC03}.

This survey is organized as follows: We give a basic classification of BiB games in Section \ref{sec:classification}, to provide some overview of the different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}. In the remaining sections of this paper we will consider the model given in Section \ref{sec:model}, which can be classified as a sequential, finite and dynamic process with random deletions. Section \ref{sec:multiple-choice} starts with a brief definition of three different classes of random placement strategies and then continues with the presentation of the two placement algorithms outlines in the introduction: The  Greedy algorithm (Section \ref{sec:greedy}) and the Always-go-left algorithm (Section \ref{sec:AlwaysGoLeft}). Section \ref{sec:analysis} is the core of this survey and will give a detailed analysis of both strategies. We will use a proof technique called witness trees, which allows for a clean analysis of both problems and is applicable to a wide variety of BiB scenarios. Other proof techniques such as layered induction and a differential equations approach have been used. See, e.g., \cite{MRS01} for an overview of different proof techniques for the BiB problem.

@@

\section{Classification}
\label{sec:classification}
Several different variations of this problem exist, which may be classified as follows:
\begin{compactitem}
\item In a \emph{sequential} process, the balls arrive one after another. A  ball is dealt with only after the previous ball has been placed. Conversely, in a \emph{parallel} process, balls arrive in sets and the balls in one set are dealt with simultaneously \cite{ABS98}. 
\item In a \emph{static} problem, the balls are placed into the bins and no balls are ever removed whereas in a \emph{dynamic} problem balls can also be removed from the system. In the random deletions model, for each newly inserted ball one is chosen uniformly and independently at random and removed from the system \cite{ABKU99} \cite{MRS01}. More elaborate schemes exists: An oblivious adversary specifies the sequence of insertions and deletions of balls in advance, without any knowledge of their placement \cite{CFM+98}. Another approach was studied by Adler et al.~\cite{ABS98} in the context of the parallel BiB problem: In each bin, the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted. 
\item A \emph{finite} process is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite} process, a placement and a deletion process is defined over an infinite time span \cite{ABS98}.
\item In the lightly loaded case, $m$ is in the order of $n$, whereas in the heavily loaded case, $m$ is arbitrarily large. We will assume that $m = n$ in the lightly loaded case. 
\end{compactitem}

\section{Applications}
\label{sec:applications}
The two-choice paradigm applied to the BiB model has several useful applications, some of which we will outline here. Hashing (Section \ref{sec:hashing}), online load balancing (Section \ref{sec:loadbalancing}) and the emulation of shared memory machines on distributed memory machines \ref{sec:DMM}) share a close relation to load balancing. Section \ref{sec:circuitrouting} applies the BiB model and the two-choice paradigm in a different context, namely that of routing circuits in multi-stage interconnection networks with minimal congestion.

@@

It is easy to see how the BiB idea can be applied: Assuming a perfect hash function\footnote{i.e., each element is mapped to one of the $n$ table entries uniformly and independently at random}, the elements are balls to be mapped to table entries or bins, implying a worst case look-up time of $\Theta\left(\frac{\log n}{\log \log n}\right)$ w.h.p. Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries implies a look-up time of $\Theta\left(\log \log n\right)$ w.h.p. This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions \cite{ABKU99} \cite{MRS01}.

\subsection{Online load balancing}
\label{sec:loadbalancing}
The following scenario is a typical application of online load balancing: Consider $n$ servers (such as file-, database- or network servers) and $m$ requests. For simplicity, assume that each task can be handled by any server and that all tasks are of the same size, i.e., they require the same execution time.

To maximize the throughput of the servers, a uniform distribution of tasks to servers is desirable. While a central dispatcher dispatcher can easily achieve this, it also poses a bottleneck within a distributed system since querying $m$ servers for their current load involves significant overhead (e.g., sending a message to each server and waiting for the reply). As with hashing, the BiB model can be applied easily: we need to assign tasks (the balls) to servers (the bins). Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most$O\left(\log \log n\right)$ w.h.p. \cite{KLM92} \cite{RS98} \cite{MRS01}.

\subsection{Emulation of shared memory machines on distributed memory machines}
\label{sec:DMM}
The Parallel Random Access Machine model (PRAM) ~\cite{P03} is a high level abstraction of a shared memory machine, in which memory access by any processor to any memory location is assumed to require constant time. It is especially useful for the study of parallel algorithms. However, a PRAM is much harder to realize in hardware than a Distributed Memory Machine (DMM). Thus, ways to emulate a PRAM on a DMM have been studied extensively. Such an emulation distributes the processors and memory cells of the PRAM to the processors and memory modules (typically one per processor) of the DMM with the goal to reduce the slowdown of the emulation, which is the time the DMM requires to simulate one step of the PRAM. Each memory module can handle only one memory access at a time step. More accesses to the same module will result in memory contention. The BiB idea is useful for managing communication between the processors and the memory modules such that memory contention is minimal. For details, see Karp et al.~\cite{KLM92} and the papers referenced there.

\subsection{Low congestion circuit routing}
\label{sec:circuitrouting}
In multi-stage interconnection networks, virtual circuit-switching is used to route communication. For simplicity, we consider the permutation routing problem, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and assume a butterfly network as the underlying interconnection network~\cite{CLR09}. Then, the congestion is the maximum number of paths that are routed through a network link. The goal of a circuit routing algorithm is to to allocate paths for all $n$ communication requests such that congestion is minimized.

@@

Applying the two-choice paradigm in this context is somewhat more involved. The interested reader may find more details in \cite{CMM+98}.

@@

Consider the Greedy scheme, a variation of the one-choice BiB strategy: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin. If there are multiple bins with the same minimal load, chose one among them arbitrarily. Note that the $d$ selected bins are not necessarily distinct.

\begin{theorem}
\label{theorem:greedy}
Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is bounded as 
\begin{align*}
\load \leq \frac{m}{n} + \frac{\log \log n}{\log d}+ O(h)
\end{align*}
w.h.p.~for all $t$ \cite{ABKU99} \cite{BCSV06}.
\end{theorem}

This implies a gap of $O\left(\log \log n \right)$ for arbitrarily large $m$, which is remarkable in several ways:
\begin{compactitem}
\item The apparently small change made for the Greedy algorithm compared to the single-choice BiB scheme results in an exponential decrease of the gap, even for $d=2$. 
\item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{MRS01}. 
\item For the heavily loaded case, the resulting gap of the multiple-choice BiB strategy does not depend on the number of balls $m$. In contrast, the bound for the one-choice BiB scheme diverges with $m$.
\begin{comment}
\todo{the following is only true for class 1 algorithms}
\item The given bounds are tight, meaning that no other strategy that places each ball into one of $d$ randomly selected bins achieves a gap that is asymptotically lower.
\end{comment}
\end{compactitem}

@@

\begin{theorem}
\label{theorem:agln}
If $n$ balls are placed into $n$ bins using the Always-go-left scheme, the maximum load will be
\begin{align*}
\load \leq \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
\end{align*}
w.h.p.
\end{theorem}

\begin{theorem}
\label{theorem:algm}
Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, at any time $t$ the maximum load is at most 
\begin{align*}
\load \leq \frac{\ln\ln n}{d  \ln \Phi_d} + O(h)
\end{align*}
w.h.p.
\end{theorem}

\section{Analysis}
\label{sec:analysis}
In this section, we will prove Theorems \ref{theorem:greedy}, \ref{theorem:agln} and \ref{theorem:algm} in detail, using witness trees to provide an upper bound for the event that a bin contains too many balls. Our analysis closely follows the proofs in \cite{VOC03}.

@@

We define edge and leaf events:
\begin{compactitem}
\item \emph{Edge event:} Consider some edge $e = (u,v)$ of the witness tree, s.t.~$v$ is the $i$-th child of $u$. Then $e$ represents the event that the $i$-th location of the ball $b_u = \mathrm{ball}(u)$ points to the same bin as one of the locations of the ball $b_v = \mathrm{ball}(v)$. 
%$\todo{illustrate}
\item \emph{Leaf event:} A leaf node $v$ of the witness tree represents the event that each of the $d$ locations of the ball $b_v = \mathrm{ball}(v)$ points to a bin that contains at least three additional balls (i.e., balls that are not represented by a node of the tree) at the time of the insertion of the ball $b_v$.
%\todo{illustrate}
\end{compactitem}
The definition for leaf events refers to bins containing some number of balls at some specified time. To avoid dependencies, the crucial trick is that edge events are defined in terms of the alternative locations of balls, not by their finally assigned bin.

@@

An edge\footnote{where $v$ is the $i$-th child of $u$} $(u, v)$ is activated with probability at most $d/n$, which is the probability that the $i$-th location of the ball $b_u = \mathrm{ball}(u)$ points to a particular location of the ball $b_v = \mathrm{ball}(v)$ times the number of locations $d$. This implies a probability of at most
\begin{align*}
p_e = \left(\frac{d}{n}\right)^{k-1}
\end{align*} 
or the event that all $k-1$ edges of a witness tree are activated. Note that  the assumption that all balls are distinct (Assumption \ref{assumption:independence}) implies that all edge events are independent.

@@

Let $x$ be the bin that contains $L+4$ balls. The root node gets assigned the topmost ball in bin $x$. Each of the $d$ locations of this ball points to a bin with at least $L+3$ balls in it, at the time of its insertion. These are the $d$ bins that were sampled by the Always-go-left scheme and thus bin $i$ belongs to the $i$'th group of bins, where each group consists of $n/d$ bins\footnote{For the sake of simplicity, we assume that $d$ divides $n$ evenly, i.e., that all groups have the exact same size. It can be shown that the analysis still holds if each group is of size $\lfloor\frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$.}. The child node of the root with label $(L-1, i)$ is assigned the topmost ball of the bin the $i$' location points to. The assignment of balls to nodes than proceeds recursively. Given a node $v$ with label $(h, i)$ whose ball $b_v$ was put into some bin belonging to group $i$ and containing at least $l+3$ other balls, we can conclude the following.
\begin{compactitem}
\item For $1\leq j < i$, the $j$'th location of the ball $b_v$ points to a bin containing at least $l+4$ balls at the insertion time of the ball $b_v$. If the $j$-th location contained less balls than the $i$-th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$. The topmost ball in location $j$ is assigned to the  child $u$ of $v$ with label $(l,j)$. 
\item For $i < j \leq d$, the $j$-th location contains at least $l+3$ balls. If location $j$ contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin it points to. The topmost ball in the bin that location $j$ points to is assigned to the child node $u$ of $v$ labeled with $(l-1, j)$. \todo{the paper claims that the locations $j$ point to bins with $l+2$ balls. This can not be true, since the topmost, i.e., the $l+2$ ball - call it $b_t$ - gets assigned to a node with label $(l'=l-1, j)$. Since there are $l+1=l'+2$ balls below it, at the time of the placement of ball $b_t$, the bin $j$ it gets put into has only $l'+2$ balls in it, violating the invariant. Same with the case $j<i$.} 
\end{compactitem}
\end{proof}

@@

\subsubsection{Construction}
\label{sec:analysis:constructionFullWT}
Assume a bin $x$ contains at least $L+3+\gamma$ balls at time $t$. We construct a witness tree as follows.
\begin{compactitem}
\item The root is not assigned any ball.
\item Each child of the root is assigned one of the topmost $\gamma$ balls in bin $x$, which are necessarily distinct.
\item For the assignment of nodes to the grandchildren of the root, consider a ball $b_c$ that is assigned to a child node $v_c$. Let the child node of $v_c$ be denoted by $v_g$. Since the ball $b$ was placed into bin $x$, at least one of its locations points to this bin $x$, which we call location $i$. Let $j = i+1 mod d$ be the ball's $j$-th location. Then, at the time of insertion of ball $b_c$, its $j$-th location points to a bin $y$ with at least $L+3$ other balls in it. We assign the topmost ball in bin $y$ to node $v_g$.
\end{compactitem}
Notice that the $\gamma$ additional balls in bin $x$ are assigned to the child nodes of the root only.
